---
title: "p8130_homework1"
author: "Lauren Holley"
date: "2026-09-28"
output: github_document
---

```{r message=FALSE}
library(tidyverse)
```

## Question 8

#### Part A

```{r}
d = read.csv("homework1_clinic.csv",
             colClasses = c(id = "character"))

with(d, table(program, improved))
```

```{r}
d |>
  group_by(program) |>
  summarise(
    total = n(),
    observed_improved = sum(!is.na(improved)),
    number_improved = sum(improved, na.rm = TRUE),
    improvement_proportion = mean(improved, na.rm = TRUE),
    missing_waits = sum(is.na(wait_min))
  )
```

#### Part B

```{r}
rate = tapply(d$improved,
               d$program, mean)

rate

rate["Yes"] - rate["No"]
```

The Yes minus No difference in improvement proportions is 0.3333 or 33.33 percentage points. 

#### Part C

```{r improvement_plot}
barplot(rate, col = "#8DCFEC",
        ylim = c(0, 1),
        xlab = "Program Group",
        ylab = "Improvement Proportion")
```

#### Part D

Those in the program group had a higher improvement proportion, however the difference found is not automatically causal because program enrollment was voluntary. One possible confounder is that participants who are experiencing more severe symptoms could be more likely to seek program enrollment, impacting baseline measurements. Furthermore, if they begin the program with more severe symptoms, this would impact further down the line improvement levels. 

## Question 10

#### Part A

Without replacement:

P(all 3 allergy) = (3/8)(2/7)(1/6) = 0.0179

With replacement:

P(all 3 allergy) = (3/8)(3/8)(3/8) = 0.0527

#### Part B

Without replacement: 

```{r}
set.seed(813110)
B = 100000

draws = replicate(B,
  sample.int(8, 3, replace = FALSE))

all3 = colSums(draws <= 3) == 3

c(theory = (3/8)*(2/7)*(1/6),
  empirical = mean(all3))

sum(all3)
```

With replacement:

```{r}
set.seed(813111)

draws = replicate(B,
  sample.int(8, 3, replace = TRUE))

all3 = colSums(draws <= 3) == 3

c(theory = (3/8)*(3/8)*(3/8),
  empirical = mean(all3))

sum(all3)
```

Without replacement:

* Successes: 1,792

* Empirical probability: 0.0179

* Theoretical probability: 0.0179

With replacement:

* Successes: 5,213

* Empirical probability: 0.0521

* Theoretical probability: 0.0527

#### Part C

```{r}
running = cumsum(all3)/seq_len(B)

running[c(100, 1000, 10000, 100000)]
```

```{r}
plot(seq_len(B), running,
     type = "l",
     xlab = "Number of Trials",
     ylab = "Running Empirical Probability")

abline(h = 1/56)
```

#### Part 4

When there is no replacement, the probability changes as people with the allergy are selected, whereas with replacement this does not occur. The law of large numbers does not guarantee exact agreement after 25 trials, as there can always be variation. However an increase in trials does lead to running probability closer to the expected. 
